Implementing \Kwd{DNS} in \Kwd{C0}

### 2017-12-13 paper.tex
Updated networking section.

In its current state, \Kwd{C0} has a very limited standard library. In particular, while there are libraries for input and output to files and to console, there is no networking library. It may be possible to implement networking using file I/O, but we do not believe that this is reasonable.

As such, we will write a networking library for \Kwd{C0}. Due to limitations of the language, however, this library cannot be written entirely in \Kwd{C0} itself. Instead, we will write a collection of small functions in \Kwd{C} which act as an interface between \Kwd{C0} code and the \Kwd{C} sockets library. On top of this, we can then write the \Kwd{C0} portion of the library, which will provide a higher-level interface for network communication than that provided by the \Kwd{C} portion of the library.

By splitting the library into these two parts, we can keep the \Kwd{C} portion of our codebase limited in scope. In particular, we plan to ensure that we never dynamically allocate memory from \Kwd{C} and that any accesses to \Kwd{C0}-allocated memory are done so safely. The former is easily checked for, simply by checking that there are no calls to malloc or similar functions, while the latter can be achieved (but less easily checked) by using a collection of functions for safe memory access provided to libraries by the \Kwd{C0} runtime. As a result of this and the small size of the \Kwd{C} portion of the codebase, we can easily check all of the \Kwd{C} code for memory safety. Combining this with \Kwd{C0}'s memory safety, we ensure that any memory safety violations occur outside of our project (e.g. in the \Kwd{C} sockets library or in the OS).

Since this project is limited in scope to \Kwd{DNS}, we will not expose the full functionality of the sockets library in our network interface. We will instead focus on the functions that create and work with datagram-based sockets, as most of the network communication needed for \Kwd{DNS} is done via the datagram protocol \Kwd{UDP}. Additionally, to avoid busy-waiting, we will provide a limited interface to the \texttt{poll} function of the \Kwd{C} standard library in much the same way. We believe that this is sufficient to implement our nameserver, but additional library functionality can be added using these methods if necessary.

@@

By virtue of having written the majority of our \Kwd{DNS} server in
the memory- and type-safe \Kwd{C0} language, we enjoy certain safety
guarantees that are either impossible or intractable to achieve for
mainstream \Kwd{C}-based implementations. In return, we have sacrified
some performance, but it is hard to tell how much of our performance
trade-off was due to garbage collection vs.\ the unoptimized character of
our implementation (but see Section~\ref{sec:performance}).

>>>>

We have implemented a (limited) networking library for \Kwd{C0}. Due to limitations of the language, however, this library is not written entirely in \Kwd{C0} itself. Instead, we wrote a collection of small functions in \Kwd{C} which act as an interface between \Kwd{C0} code and the \Kwd{C} sockets library.

By splitting the library into these two parts, we can keep the \Kwd{C} portion of our codebase limited in scope. While our original plan was to never dynamically allocate memory in this code, we found that it was infeasible to use the socket functions without either dynamically allocating memory or violating other portions of our safety guarantees. As such, dynamically allocating memory was the safer choice, and so we have one allocation in each of the \texttt{sendto} and \texttt{recvfrom} functions. These are the only potentially unsafe regions of memory used in the application, and so we closely hand-checked their use for correctness. Additionally, we were able to ensure that any accesses to \Kwd{C0}-allocated memory are done so  using a collection of functions for safe memory access provided to libraries by the \Kwd{C0} runtime, ensuring safety of these accesses. As a result of this and the small size of the \Kwd{C} portion of the codebase, we can easily check all of the \Kwd{C} code for memory safety. Combining this with \Kwd{C0}'s memory safety, we ensure that any memory safety violations occur outside of our project (e.g. in the \Kwd{C} sockets library or in the OS).

Since this project is limited in scope to \Kwd{DNS}, we do not expose the full functionality of the sockets library in our network interface. We implemented functions that create and work with datagram-based sockets, as most of the network communication needed for \Kwd{DNS} is done via the datagram protocol \Kwd{UDP}. Additionally, to avoid busy-waiting, we provide a limited interface to the \texttt{poll} function of the \Kwd{C} standard library in much the same way. Finally, we add \Kwd{C0} interfaces for the collection of functions that translate between network and host byte order, in order to shift as much work as possible into \Kwd{C0}, rather than changing byte order in our \Kwd{C} code.

@@

By virtue of having written the majority of our \Kwd{DNS} server in
the memory- and type-safe \Kwd{C0} language, we enjoy certain safety
guarantees that are either impossible or intractable to achieve for
mainstream \Kwd{C}-based implementations. In return, we have sacrificed
some performance, but it is hard to tell how much of our performance
trade-off was due to garbage collection vs.\ the unoptimized character of
our implementation (but see Section~\ref{sec:performance}).
